### 2018-07-12 source/Ch2-techniques/imaging.tex
added various bits that came about after reshuffling the organisation

@@

\pg
Astronomers who request observation time on interferometric arrays are generally not experts in the theory and operation of these arrays. This is a simple consequence of scientific division of labour. This thesis is written from the perspective of an ``expert" user: one who specialises in understanding interferometric arrays and reducing their data, rather than their subsequent analysis. However, it can be very helpful to contextualise this perspective in the experience of other scientists, which is what this section aims to do.

\pg
LOFAR data can be acquired through observation or via the Long-Term Archive. The worst of this data will already be flagged. Users can thus generally proceed directly to calibration. Because there is no guarantee of sufficient signal-to-noise in the ``target field" (the object that the astronomer is actually interested in), it is standard practice to observe a calibrator source for 5 minutes before and after each observation. Such sources typically need to be bright and unresolved for the interferometer being used. Because they are bright and at phase centre for the 5-minute observation, the SNR when solving for calibration solutions for these 5 minutes will tend to be quite good. These solutions will then be interpolated between the 5 minutes onto the 8-hour observation. Calibration will be described in \cref{section.calibration}.

@@

\pg
Interferometric observations follow a very familiar pattern: the elements of the interferometric array (which can be single dishes or phased arrays\footnote{Described in \cref{sec.visibility}} of smaller antennas) measure a voltage. This voltage is correlated between antennas, as will be described shortly. Each of these correlations correspond to a single Fourier mode of the sky, as per the Zernike-van Cittert theorem (cf. \cref{sec.imag.psf}). They must be calibrated, by comparing the observations of a calibrator source and modelled visibilities. Calibration is described in much greater detail in \cref{section.RIME}. Then, the calibrated visibilities are used to reconstruct an image of the sky, which often requires deconvolution due to the sparsity of interferometric arrays.

@@

\pg
We have defined what a baseline corresponds to: a vector coordinate in $(u,v,w)$-space. To each baseline we associate a measurement, which we call the \emph{visibility}. 
\begin{figure}[ht]
\centering
\includegraphics[width=.7\textwidth]{images/visibility-creation.png}
\caption{\label{fig.visibility} There are two ways to combine the voltages from two antennas into a visibility: they are sum-correlation and $\pi$-correlation. In this manuscript, we will only concern ourselves with the latter. Image credit: Julien Girard}
\end{figure}
The visibility associated with baseline $\mathbf{b}_{AB}$ is created by taking the voltage measured by antenna A, multiply it by the complex conjugate of the voltage measured by antenna B, average over the correlator dump time (i.e. the time over which the measurement is made). This scalar quantity is associated to the baseline position vector $\mathbf{b}_{AB}$. In other words:
\begin{align}
\mathbf{b}_{AB} &= \frac{\mathbf{x}_{B}-\mathbf{x}_{A}}{\lambda_\mathrm{obs}}\\
V_{AB}          &= \langle V_{A} V_{B}^*\rangle_{\delta t, \delta \nu} 
\end{align}
where $\langle \cdots \rangle_{i}$ denotes an averaging over quantity $i$. We see that a visibility is a complex vector quantity. We also see that $\mathbf{b}_{AB} = \mathbf{b}_{BA}^*$: the information of the visibility associated with baseline BA is contained in the visibility associated with baseline AB. This means that in practice, only half of the visibilities ever need be stored. What does the visibility measurement correspond to?
\begin{figure}[ht]
\centering
\includegraphics[width=.5\textwidth]{images/visibility-measure.png}
\caption{\label{fig.visibility.measure} Here, we assume that there are only two sources in the sky whose signal can be measured by our antennas. The final visibility is the sum of the visibilities associated with each individual source. Image credit: Julien Girard \textcolor{red}{add reference to julien's phd}}
\end{figure}

\pg
Different astrophysical sources are located very far apart, and so their signals do not interfere. They are measured additively in both antennas. Provided that the signal from both sources is coherent when observed by the dishes, the correlation between the voltages measured by antennas A and B will simply be the sum of the voltage correlations associated with individual sources. This is shown in \cref{fig.visibility}: the radio signal from two different sources, at two different locations, is picked up simultaneously by both radio antennas. Because only the correlation between both antennas is measured, the signals do not interfere. % - i.e. the interferometric signal from different sources are additive.
\pg
Note that in Fig. \ref{fig.visibility.measure}, neither source is at the zenith. This introduces the notion of the \emph{effective baseline} which is shorter than the \emph{physical baseline} when measuring a signal from a source that is not at the zenith. Sources in different positions in the sky will be measured by an interferometric array, but the same array will have different effective baselines for different sources. To point an interferometric array in a specific direction in the sky, one must introduce a phase delay in each antenna (or, for fundamental interferometer elements such as the LOFAR tiles or NenuFAR, by playing with the cable length between each dish and the correlator). This ensures that the constituent antennas' sensitivities are maximised in the direction of interest. The position of sources are then given in terms of the cardinal angle $\sigma$: this is the difference between the position of a source and the phase centre, which is where the array is pointing. For example, in \cref{fig.visibility.measure}, if we point antennas and introduce phase delays such that we are pointing the array towards Source 2, the position of Source 1 will be $\sigma$.

@@

\pg
Assuming that all the antennas in an array are equivalent and perfectly calibrated, the van Cittert-Zernike theorem (\citetads{1934Phy.....1..201V}, covered in \citetads{2001isra.book.....T}) allows us to equate a visibility with a single Fourier mode of the plane tangent to the sky where the instrument is pointed. This direction is called the \emph{phase centre}, because phase shifts are introduced between antenna voltages before averaging so as to point each visibility in this direction.

\pg
The Point-Spread Function (PSF) of an instrument is its point source response: it will determine the angular resolution limit due to the instrument. We know that the Fourier transform of a single point source with unit flux at phase centre is a constant. If we measured the full (infinite) visibility space, then we could reconstitute a point source perfectly. In practice, however, we only ever measure a subset of the visibility space: an interferometric array's PSF is thus the inverse Fourier transform of the array's $uv$-coverage, as these are the only points for which we have information on visibility space. The more elements in the array, the better the $uv$-coverage and thus the better the PSF. An interferometer's PSF will always be worse than an antenna dish of equivalent diameter, the UV-coverage of which is the inverse Fourier transform of a disc. This is because an interferometer behaves like a \emph{sparse} dish: the sparsity of our coverage manifests as very large point-spread function. The more baselines are available, the less sparse the UV-coverage, and so the better the PSF. This is shown in Fig. \ref{fig.vla.PSFvsUVcov}.
\begin{figure}[h!]
\centering
\resizebox{\hsize}{!}{\includegraphics{images/PSFvsUVcov.png}}
\caption{\label{fig.vla.PSFvsUVcov}Plot of the PSF associated to more and more complete UV-coverage (here, consisting of more and more elements of the VLA). Image credit: Rick Perley, 3GC4 lecture.}
\end{figure}

@@

\pg
We use dynamic range (DR) as a quality metric in radio images. DR is the ratio of the flux of the brightest source in the field to the flux of the faintest source in the field.  We define dynamic range as:
\begin{equation}\label{eq.DR.imag}
\mathrm{DR} = \frac{\mathrm{flux}_\mathrm{brightest}}{\mathrm{flux}_\mathrm{faintest}}
\end{equation}
The higher the DR, the deeper we image the sky, and thus the better the image. There are two big constraints to reckon with: firstly, we do not want to deconvolve the PSF from all pixels in the image, but only from those within which large amounts of underlying flux fall. Secondly, if we only deconvolve some pixels rather than performing a full deconvolution, then care must be taken not to treat artefacts in the field (which can be created from overlapping PSF sidelobes from neighbouring sources, even with perfect calibration) as true flux. Doing so would lead to an attempt to deconvolve a PSF from a pixel while assuming an incorrect underlying flux value for this pixel and thus result in the introduction of further artefacts in the image.

\pg
The dominant family of deconvolution algorithms used in radio astronomy at the time of writing is CLEAN\footnote{The other main family of deconvolution algorithms, Maximum Entropy Methods (MEM), are not very widely-used at the time of writing, but remain an active area of research (cf. \citetads{2018ApJ...857...23C}).}. It attempts to optimise for the first requirement while being constrained by the second.The various CLEAN algorithms seek out the brightest pixel(s) in the image, potentially applying a mask to the image first in cases where some \emph{a priori} knowledge on flux distribution is available. It then deconvolves those pixels up to a predefined threshold (either a fraction of the initial pixel value or some factor of the estimated noise value). It does this some predefined number of times. Then, it collects the aggregate \emph{model} (the flux estimate for each deconvolved pixel up to this point), convolves the model with the PSF, and subtracts the result from the dirty image. It then begins to choose the brightest pixels again, and continues to clean until it reaches a stopping condition (usually either a flux value or a maximum number of iterations).

\begin{figure}[h!]
\centering
\resizebox{\hsize}{!}{\includegraphics{images/{dirty-vs-clean}.png}}
\caption{\label{fig.vla.clean} Effect of MetroCLEAN algorithm on a dirty image of 3C295. To the left, the dirty image; to the right, the deconvolved image (same size, resolution \& flux scale, made with the same data). As we can see, the latter is much more scientifically interesting.}
\end{figure}

\pg
In Fig. \ref{fig.vla.clean}, we show the effect of deconvolution on a dirty image of 3C295 (a very bright \& complex source which is the subject of much of this thesis' work). Since this source includes diffuse emission, standard CLEAN is insufficient (since CLEAN assumes a sparse distribution of point-like sources). Instead, the DDFacet software package's novel MetroCLEAN algorithm (see \citetads{2017arXiv171202078T}), which deconvolves islands consisting of sets of pixels at once, was used. This allowed for a far greater dynamic range than standard CLEAN could achieve.

@@

\section{Weighting Schemes}
\pg
In \cref{section.clean}, we discussed the problem of deconvolution in radio interferometric imaging. One consideration that is not raised is the issue of conditioning. Consider two extreme cases: a field containing a few point sources(\cref{fig.point}), and a field containing a diffuse and turbulent source, with very complex flux distribution at all scales (\cref{fig.cyga}). Deconvolving the PSF from the first image using CLEAN algorithms will be child's play, but doing the same with the second will be extremely complex. The first case is said to be well-conditioned, and the second to be poorly-conditioned.

@@

\pg
We close this section by noting that data flagging can be considered a form of weighting scheme. This consists of assigning a null weight to those data points considered too corrupted by various instrumental or atmospheric effects (typically, Radio Frequency Interference - RFI - cf. \citetads[cf.]{2010ascl.soft10017O} for an example) to be scientifically useful. Those weights which are not deemed too corrupted are assigned a unit weight. This corresponds to simply dropping those data points which would corrupt the image reconstruction by introducing unphysically strong fringes in the image.

>>>>

\pg
One way to move past this technical limitation is to resort to interferometry: this is the topic of this section.

@@

\pg
Astronomers who request observation time on interferometric arrays are generally not experts in the theory and operation of these arrays. This is a simple consequence of scientific division of labour. This thesis is written from the perspective of an ``expert" user: one who specialises in understanding interferometric arrays and reducing their data, rather than their subsequent analysis. However, it can be very helpful to contextualise this perspective in the experience of other scientists, which is what this section aims to do. This is simply to highlight the conceptual structure of a radio interferometric observation, and where the concepts outlined in this chapter fall in the overall pattern. In other words, it is best understood as a guide to see where a given section's topic slots into the overall production of an image made from visibilities.
\pg
LOFAR data can be acquired through observation or via the Long-Term Archive. The worst of this data will already be flagged. Users can thus generally proceed directly to calibration. Because there is no guarantee of sufficient signal-to-noise in the ``target field" (the object that the astronomer is actually interested in), it is standard practice to observe a calibrator source for 5 minutes before and after each observation. Such sources typically need to be bright and unresolved for the interferometer being used. Because they are bright and at phase centre for the 5-minute observation, the SNR when solving for calibration solutions for these 5 minutes will tend to be quite good. These solutions will then be interpolated between the 5 minutes onto the 8-hour observation. Calibration will be described in \cref{section.calibration}.

@@

\pg
Interferometric measurements are made as follows: the elements of the interferometric array (which can be single dishes or phased arrays\footnote{Described in \cref{sec.visibility}} of smaller antennas) measure a voltage. This voltage is correlated between antennas, as will be described shortly. Each of these correlations correspond to a single Fourier mode of the sky, as per the Zernike-van Cittert theorem (cf. \cref{sec.imag.psf}). They must be calibrated, by comparing the observations of a calibrator source and modelled visibilities. Calibration is described in much greater detail in \cref{section.RIME}. Then, the calibrated visibilities are used to reconstruct an image of the sky, which often requires deconvolution due to the sparsity of interferometric arrays.

@@

\pg
We have defined what a baseline corresponds to: a vector coordinate in $(u,v,w)$-space. To each baseline we associate a measurement, which we call the \emph{visibility}. 
\begin{figure}[ht]
\centering
\includegraphics[width=.7\textwidth]{images/visibility-creation.png}
\caption{\label{fig.visibility} There are two ways to combine the voltages from two antennas into a visibility: they are sum-correlation and $\pi$-correlation. In this manuscript, we will only concern ourselves with the latter. Image credit: Julien Girard}
\end{figure}
The visibility associated with baseline $\mathbf{b}_{AB}$ is created by taking the voltage $V_A$ measured by antenna A, multiply it by the complex conjugate $V_B^*$ of the voltage measured by antenna B, average over the correlator dump time (i.e. the time over which the measurement is made). This scalar quantity is associated to the baseline position vector $\mathbf{b}_{AB}$. In other words:
\begin{align}
\mathbf{b}_{AB} &= \frac{\mathbf{x}_{B}-\mathbf{x}_{A}}{\lambda_\mathrm{obs}}\\
V_{AB}          &= \langle V_{A} V_{B}^*\rangle_{\delta t, \delta \nu} 
\end{align}
where $\langle \cdots \rangle_{i}$ denotes an averaging over quantity $i$. We see that a visibility is a complex vector quantity. We also see that $\mathbf{b}_{AB} = \mathbf{b}_{BA}^*$: the information of the visibility associated with baseline BA is contained in the visibility associated with baseline AB. This means that in practice, only half of the visibilities ever need be stored. What does the visibility measurement correspond to?
\begin{figure}[ht]
\centering
\includegraphics[width=.5\textwidth]{images/visibility-measure.png}
\caption{\label{fig.visibility.measure} Here, we assume that there are only two sources in the sky whose signal can be measured by our antennas. The final visibility is the sum of the visibilities associated with each individual source. Image credit: Julien Girard \textcolor{red}{add reference to julien's phd}}
\end{figure}

\pg
Different astrophysical sources are located very far apart, and so their signals do not interfere. They are measured additively in both antennas. Provided that the signal from both sources is coherent when observed by the dishes, the correlation between the voltages measured by antennas A and B will simply be the sum of the voltage correlations associated with individual sources. This is shown in \cref{fig.visibility}: the radio signal from two different sources, at two different locations, is picked up simultaneously by both radio antennas. Because only the correlation between both antennas is measured, the signals do not interfere: if the signal from each source gives a voltage $V_{p}^s$ for antenna $p$, then the additive property of electric fields means that the total visibility measured by baseline $pq$ can be written as:
\begin{align}
V_{pq}  &= \langle \left(\sum_{s} V_{p}^s\right) \left(\sum_{s} (V_{q}^s)\right)^*  \rangle_{\Delta t} \\
        &= \langle \sum_{s,s'} V_{p}^s (V_q^{s'})^*  \rangle_{\Delta t}\\
        &= \langle \sum_s V_{p}^s (V_q^{s})^* \rangle_{\Delta t} + \underbrace{\langle \sum_{s,s'\ne s} V_{p}^s (V_q^{s'})^* \rangle_{\Delta t}}_{=0} \label{eq.spatial.incor}
\end{align}
where the fact that signal from different sources is not coherent means that we average out cross-contributions (i.e. signal from $s\ne s'$). 
%\begin{align}
%V_{pq}&=\langle \left(\sum_{s} V_{p}^s\right) \left(\sum_{s} V_{q}^s \right)^* \rangle_{\Delta t} \right)\\
%      &=\langle\sum_{s,s'} \left( V_p^s\left(V_q^{s'}\right)^* \rangle_{\Delta t}
%\end{align}

\pg
Note that in Fig. \ref{fig.visibility.measure}, neither source is at the zenith. This introduces the notion of the \emph{effective baseline} which is shorter than the \emph{physical baseline} when measuring a signal from a source that is not at the zenith. Sources in different positions in the sky will be measured by an interferometric array, but the same array will have different effective baselines for different sources. To point an interferometric array in a specific direction in the sky, one must introduce a phase delay in each antenna (or, for fundamental interferometer elements such as the LOFAR tiles or NenuFAR, by playing with the cable length between each dish and the correlator). Because the phase needs to be properly sampled to limit information loss, a process known as fringe stopping\footnote{see \href{https://casper.berkeley.edu/astrobaki/index.php/Fringe_Stopping}{here} for more information. Fringe stopping consists of measuring the beat between a predicted fringe rate (i.e. the period of the visibility sine wave corresponding to a point source out of phase centre; if the position of an array's antennas is well-known, this can be modeled accurately) and the actual fringe rate (i.e. the fringe rate corresponding to the true period of the visibility sine wave). The difference between the two is a function of how well the antennas' positions are known. If they are well-known, then the phase information can still be measured, even without very fine sampling in time. } must be used to avoid taking a measurement every few milliseconds. The position of sources are then given in terms of the cardinal angle $\sigma$: this is the difference between the position of a source and the phase centre, which is where the array is pointing. For example, in \cref{fig.visibility.measure}, if we point antennas and introduce phase delays such that we are pointing the array towards Source 2, the position of Source 1 will be $\sigma$.

@@

\pg
Assuming that all the antennas in an array are equivalent and perfectly calibrated, the van Cittert-Zernike theorem (\citetads{1934Phy.....1..201V}, covered in \citetads{2001isra.book.....T}) allows us to equate a visibility with a single Fourier mode of the plane tangent to the sky where the instrument is pointed. This direction is called the \emph{phase centre}, because phase shifts are introduced between antenna voltages before averaging so as to point each visibility in this direction. This behaviour is encoded in the Fourier kernel, which is written as:
\begin{equation}
k_p=2\pi i (( u_p l + v_p m + w_p (n-1) ))
\end{equation}

\pg
The Point-Spread Function (PSF) of an instrument is its point source response: it will determine the angular resolution limit due to the instrument. We know that the Fourier transform of a single point source with unit flux at phase centre is a constant. If we measured the full (infinite) visibility space, then we could reconstitute a point source perfectly. In practice, however, we only ever measure a subset of the visibility space: an interferometric array's PSF is thus the inverse Fourier transform of the array's $uv$-coverage, as these are the only points for which we have information on visibility space. The more elements in the array, the better the $uv$-coverage and thus the better the PSF. An interferometer's PSF will always be worse than an antenna dish of equivalent diameter, the UV-coverage of which is the inverse Fourier transform of a disc. This is because an interferometer behaves like a \emph{sparse} dish: the sparsity of our coverage manifests as very large point-spread function. The more baselines are available, the less sparse the UV-coverage, and so the better the PSF. This is shown in Fig. \ref{fig.vla.PSFvsUVcov}.

\pg
Mathematically, the $uv$-coverage of an interferometer can be treated as a series of delta functions centred on the antenna positions. To go from visibility space to image space, we apply the inverse Fourier transform, giving us:
\begin{align}
\mathbf{S}_{\mathrm{obs}} &= \mathcal{F}^*\left[ \mathbf{B} \sum_i^{n} \mathbf{x}_i \right]
\end{align}
where $\mathbf{S}_{\mathrm{obs}}$ is the observed sky brightness distribution, $\mathbf{B}$ the Fourier transform of the true sky brightness distribution $\mathbf{S}$, and $n$ is the number of antennas in our array. $\mathcal{F}^*$ is the inverse Fourier transform. By using the convolution theorem, we immediately see that
\begin{align}
\mathbf{S}_{\mathrm{obs}} &= \underbrace{\mathcal{F}^*\left[ \mathbf{B}\right]}_{=\mathbf{S}} \star \underbrace{\mathcal{F}^*\left[\sum_i^{n}\mathbf{x}_i \right]}_{\text{the PSF}}\\
\end{align}
and that our image is the convolution of the ``true", underlying sky brightness distribution, and our array's PSF. As $n \rightarrow \infty$, the $uv$-plane fills, and the inverse Fourier transform of this unit plane becomes 1. This illustrates the description above.

\begin{figure}[h!]
\centering
\resizebox{\hsize}{!}{\includegraphics{images/PSFvsUVcov.png}}
\caption{\label{fig.vla.PSFvsUVcov}Plot of the PSF associated to more and more complete UV-coverage (here, consisting of more and more elements of the VLA). Image credit: Rick Perley, 3GC4 lecture.}
\end{figure}

@@

\pg
We use dynamic range (DR) as a quality metric in radio images. DR is the ratio of the flux of the brightest source in the field to the flux of the faintest source in the field detectable above the noise. It is therefore a function of both source brightness and noise, and so we define dynamic range as:
\begin{equation}\label{eq.DR.imag}
DR = \frac{S_{max}}{\max(S_{min},\sigma)}
\end{equation}
where $S_{max}$ is the flux of the brightest source in the image, $S_{min}$ the flux of the faintest source detectable above the noise, and $\sigma$ is the noise in the image. The higher the DR, the deeper we image the sky, and thus the better the image. There are two big constraints to reckon with: firstly, we do not want to deconvolve the PSF from all pixels in the image, but only from those within which large amounts of underlying flux fall. Secondly, if we only deconvolve some pixels rather than performing a full deconvolution, then care must be taken not to treat artefacts in the field (which can be created from overlapping PSF sidelobes from neighbouring sources, even with perfect calibration) as true flux. Doing so would lead to an attempt to deconvolve a PSF from a pixel while assuming an incorrect underlying flux value for this pixel and thus result in the introduction of further artefacts in the image.

\pg
The dominant family of deconvolution algorithms used in radio astronomy at the time of writing is CLEAN\footnote{The other main family of deconvolution algorithms, Maximum Entropy Methods (MEM), are not very widely-used at the time of writing, but remain an active area of research (cf. \citetads{2018ApJ...857...23C}).} \citepads[see]{1974A&AS...15..417H,1984ARA&A..22...97P}, which is understood in compressed sensing theory as a matching-pursuit algorithm optimising L2 (i.e. a least-squares fit) with the addition of an L1 regularisation (i.e. the total flux in the model is minimised). The various CLEAN algorithms seek out the brightest pixel(s) in the image, potentially applying a mask to the image first in cases where some \emph{a priori} knowledge on flux distribution is available. It then deconvolves those pixels up to a predefined threshold (either a fraction of the initial pixel value or some factor of the estimated noise value). It does this some predefined number of times. Then, it collects the aggregate \emph{model} (the flux estimate for each deconvolved pixel up to this point), convolves the model with the PSF, and subtracts the result from the dirty image. It then begins to choose the brightest pixels again, and continues to clean until it reaches a stopping condition (usually either a flux value or a maximum number of iterations).

\begin{figure}[h!]
\centering
\resizebox{\hsize}{!}{\includegraphics{images/{dirty-vs-clean}.png}}
\caption{\label{fig.vla.clean} Effect of deconvolution on a dirty image of 3C295. To the left, the dirty image; to the right, the deconvolved image (same size, resolution \& flux scale, made with the same data). As we can see, the latter is much more scientifically interesting.}
\end{figure}

\pg
In Fig. \ref{fig.vla.clean}, we show the effect of deconvolution on a dirty image of 3C295 (a very bright \& complex source which is the subject of much of this thesis' work). Since this source includes diffuse emission, standard CLEAN is insufficient (since CLEAN assumes a sparse distribution of point-like sources). Instead, the DDFacet software package's novel SSD/GA algorithm (see \citetads{2017arXiv171202078T}), which deconvolves islands consisting of sets of pixels at once, was used. This allowed for a far greater dynamic range than standard CLEAN could achieve, while also more properly recovering diffuse emission: when imaging a resolved source, this is absolutely necessary.

@@

\section{Unifying Calibration \& Imaging}\label{sec.CalibImagery}

\pg
The previous section mentions that good deconvolution is dependent on good calibration: poor calibration introduces artefacts in images, which can in turn bias deconvolution. A wrong model of the sky brightness distribution, in turn, will bias calibration, introducing artefacts in images. This feedback of one inverse problem onto the other is a strong indication that both are, in fact, different facets of the same underlying inverse problem: going from measured visibilities to a sky brightness distribution. Indeed, one could - conceptually, at least - solve for both problems at once: find both the sky brightness distribution and the antenna gains which best fit the measured visibilities. In practice, this is not feasible; not only is there no proof that the overall calibration \& imaging problem is convex, but even if it is, it is also poorly conditioned: multiple sets of sky brightness distributions \& antenna gains are equally viable for a given set of measured visibilities. Examples of convexity and conditioning are shown in \cref{fig.examples}.

\begin{figure}[h!]
\centering
\begin{subfigure}{.32\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/GoodConditioning.png}}
\caption{\label{fig.goodconditioning} Convex, Well-conditioned}
\end{subfigure}
\hfill
\begin{subfigure}{.32\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/BadConditioning.png}}
\caption{\label{fig.badconditioning} Convex, ill-conditioned}
\end{subfigure}
\hfill
\begin{subfigure}{.32\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/NotConvex.png}}
\caption{\label{fig.notconvex} Non-convex, ill-conditioned}
\end{subfigure}
\caption{\label{fig.examples} Examples of poor conditioning and non-convexity, plotted in terms of residual $\chi^2$ values vs parameter values.  Note that all the parameters of our problem are collapsed onto a single axis and treated as a scalar: this is for illustrative purposes. Note that the local minimum of the non-convex example is well-conditioned while its global minimum is ill-conditioned.}
\end{figure}

\pg
Because the net calibration \& imaging problem has so many degrees of freedom, it is convenient to break it into two parts: calibration and imaging. By doing this, instead of exploring a massive set of hyperparameters (e.g. relative flux distribution at a given frequency, overall spectral index, 8 values per antenna per time and frequency interval, ...), a smaller subset of these hyperparameters is explored while all else is kept constant. By ensuring that these subsets each give reasonable results individually, it becomes possible to solve the full problem. This practice of imposing specific priors to each subset of the problem is a form of regularisation: it can turn a poorly-conditioned problem into a better-conditioned one, provided that the priors are well-chosen.

\pg
In effect, this means going from exploring the sky-gains parameter space all at once to exploring it along one ``axis\footnote{which includes all the hyperparameters associated with that subset of the problem}" at a time. This is illustrated in \cref{fig.selfcal.paramexplore}. 
\begin{figure}[h!]
\includegraphics[width=.6\linewidth]{images/Selfcal-paramExplore.png}
\caption{\label{fig.selfcal.paramexplore} Plot showing the exploration of sky-gains (X-G) parameter space. In red, this space is explored one axis at a time (this is ``self-calibration"); in blue, the full space explored at once. Here, we show both procedures converging to the best confidence value: in practice, this is rarely the case.}
\end{figure}

\pg
In this context, a good deconvolution algorithm is one that is able to minimise how much it is affected by calibration errors while maximising how much true underlying flux it can recover into a model, while a good calibration algorithm is one that is able to minimise how much it is affected by sky model errors while maximising how much true underlying gain structure it can recover.

\section{Weighting Schemes}
\pg
We will finish our introduction to the inverse problem of interferometry by discussing one method by which the conditioning of deconvolution can be improved. Consider two extreme cases: a field containing a few point sources(\cref{fig.point}), and a field containing a diffuse and turbulent source, with very complex flux distribution at all scales (\cref{fig.cyga}). Deconvolving the PSF from the first image using CLEAN algorithms will be child's play, but doing the same with the second will be extremely complex. The first case is well-conditioned, and the second to ill-conditioned.

@@

\pg
We close this section by noting that data flagging can be considered a form of weighting scheme. This consists of assigning a null weight to those data points considered too corrupted by various instrumental or atmospheric effects (typically, Radio Frequency Interference - RFI - cf. \citetads[cf.]{2010ascl.soft10017O} for an example) to be scientifically useful. Those weights which are not deemed too corrupted are assigned a weight of zero. This corresponds to simply dropping those data points which would corrupt the image reconstruction by introducing unphysically strong fringes in the image.

### 2018-07-12 source/Ch2-techniques/lofar.tex
added various bits that came about after reshuffling the organisation

\pg
In this section, we will describe the LOw Frequency Array LOFAR \citepads{2013A&A...556A...2V}, its technical properties and its current state of the art. In particular, the distinction between "Dutch" LOFAR and "international" LOFAR - and the technical problems associated with each - will be made explicit in this section.

>>>>

\pg
Our observations have all been made with the LOw Frequency Array LOFAR \citepads{2013A&A...556A...2V}. In this section, we describe its technical properties and its current state of the art. In particular, the distinction between ``Dutch" LOFAR and ``international" LOFAR - and the technical problems associated with each - will be made explicit in this section.

### 2018-07-12 source/Ch2-techniques/techniques.tex
added various bits that came about after reshuffling the organisation

\chapter{Introduction}
\minitoc

\pg
In this section, we contextualise the work done as part of this PhD by discussing the difficulties introduced by the combination of sparse $uv$-coverage and weak \emph{a priori} constraints on the sky brightness distribution. The conceptual framework we will rely on throughout this manuscript, the Radio Interferometer's Measurement Equation, is described in greater details in  \cref{section.RIME}. We will begin by discussing radio interferometry itself. We will then discuss the so-called \emph{imaging problem}, and end by discussing the \emph{calibration problem}. While in practice calibration is done before imaging, it is conceptually helpful to begin with imaging. Until we reach our introduction to radio interferometric calibration, we will therefore assume that calibration has been successfully carried out, and that we are working on the \emph{corrected visibilities}, i.e. gain-corrected visibilities.

@@

% will make the ideas and basis of radio interferometry more accessible by allowing us to explain the abstractions that interferometry relies on in terms of simpler instrumental configurations. We will then introduce the concrete problems that interferometry introduces: both a recapitulation of the venerable Zernike-van Cittert theorem \citep[cf.][]{1934Phy.....1..201V} and the problem of incomplete $uv$-coverage.

\clearpage
\input{source/Ch2-intro/imaging.tex}
\clearpage
\input{source/Ch2-intro/calibration.tex}
\clearpage
\input{source/Ch2-intro/science.tex}
\clearpage
\input{source/Ch2-intro/lofar.tex}

>>>>

\chapter{Techniques of Interferometry}
\minitoc

\pg
In this section, we contextualise the work done as part of this PhD by discussing the difficulties introduced by the combination of sparse $uv$-coverage and weak \emph{a priori} constraints on the sky brightness distribution. The conceptual framework which we rely on throughout this manuscript, the Radio Interferometer's Measurement Equation, is described in greater detail in  \cref{section.RIME}. We will begin by discussing radio interferometry itself. We will then discuss the so-called \emph{imaging problem}, and end by discussing the \emph{calibration problem}. Note that both of these are artifical subsets of the general inverse problem of radio interferometry. While in practice calibration is done before imaging, it is conceptually helpful to begin with imaging. Until we reach our introduction to radio interferometric calibration, we will therefore assume that calibration has been successfully carried out, and that we are working on the \emph{corrected visibilities}, i.e. gain-corrected visibilities.

@@

% will make the ideas and basis of radio interferometry more accessible by allowing us to explain the abstractions that interferometry relies on in terms of simpler instrumental configurations. We will then introduce the concrete problems that interferometry introduces: both a recapitulation of the venerable Zernike-van Cittert theorem \citep[cf.][]{1934Phy.....1..201V} and the problem of incomplete $uv$-cove	rage.

\clearpage
\input{source/Ch2-techniques/imaging.tex}
\clearpage
\input{source/Ch2-techniques/calibration.tex}
\clearpage
\input{source/Ch2-techniques/RIME.tex}
